The Name of the Title Is Hope

### 2025-01-21 paper.tex
fixing build errors, can't have spaces or parens in citekeys

\subsection{The Scan Primitive}\label{sec:the-scan-primitive}
The study of \emph{scan} (\emph{parallel prefix}) networks traces back to the design of carry-lookahead adder circuits and beyond~\cite{5219801}. A scan is typically defined on a monoid \( M \), characterized by a binary reduction operator \( \oplus \) and an identity element \( \varepsilon \). The binary operator \( \oplus \) satisfies the closure property \( \forall a, b \in M, \ (a \oplus b) \in M \) and has an identity element \( \exists \varepsilon \in M, \ \forall a \in M, \ \varepsilon \oplus a = a \). Although \( \oplus \) must be associative, it is not necessarily commutative, as demonstrated in structures like the \emph{bicyclic semigroup}~\cite{}\footnote{Despite its name, this is a monoid.} and Rabin-Karp's \emph{second fingerprinting method for string matching}~\cite[Section 6]{Karp:1987:ERP}. In a scan, the result at the \( n \)-th element is the reduction of the preceding subset of elements in the sequence. If the subset includes the \( n \)-th element, it is called \emph{inclusive}; if it excludes the \( n \)-th element, it is called \emph{exclusive}. The most common scan type is the prefix sum, where \( \oplus \) is addition. For example:
\[
  x = [x_1, x_2, x_3, \dots, x_n] \ \ \ \ \ \ \ y = [1, 1, 1, 1, 1]
\]
\[
  \text{InclusiveScan}(x, \oplus) = [x_1, x_1 \oplus x_2, x_1 \oplus x_2 \oplus x_3, \dots, x_1 \oplus x_2 \oplus \cdots \oplus x_n]
\]
\[
  \text{InclusiveScan}(y, +) = [1, 2, 3, 4, 5]
\]
Due to its significance in circuits and as a fundamental algorithmic primitive, scan has been extensively studied in both electrical engineering and computer science. Harris~\cite{1292373} offers a taxonomy that relates depth, fanout, and wire tracks, while Hinze~\cite{10.1007/978-3-540-27764-4_11} develops an algebraic framework for scans. Snir~\cite{10.1016/0196-6774(86)90003-9} proved that depth $d$ and size $s$ are related by $s + d \ge 2n - 2$, and Fich~\cite{10.1145/800061.808738} proved that among minimum depth scans Ladner-Fischer~\cite{10.1145/322217.322232} networks have optimal size. Blelloch~\cite{} adapted scan to the PRAM model and popularized its use as an algorithmic primitive. Finally, Merrill and Garland~\cite{Merrill2016} provide a review of GPU scan implementations, complementing earlier work by Merrill and Grimshaw~\cite{Merrill2009}.

@@

But, while \emph{Chained-Scan} with \emph{Decoupled Lookback} appears to be the ideal scan architecture in terms of its theoretical and implementation characteristics, it falls short in meeting the criterion of \emph{portability}. This limitation arises because \emph{Chained-Scan} depends on a \emph{forward-progress guarantee}, a feature provided by NVIDIA devices but notably lacking on Apple's M-series and ARM GPUs~\cite{sorensen 2}. Without this guarantee, \emph{Chained-Scan} is highly prone to catastrophic failure in the form of deadlocking.

>>>>

\subsection{The Scan Primitive}\label{sec:the-scan-primitive}
The study of \emph{scan} (\emph{parallel prefix}) networks traces back to the design of carry-lookahead adder circuits and beyond~\cite{5219801}. A scan is typically defined on a monoid \( M \), characterized by a binary reduction operator \( \oplus \) and an identity element \( \varepsilon \). The binary operator \( \oplus \) satisfies the closure property \( \forall a, b \in M, \ (a \oplus b) \in M \) and has an identity element \( \exists \varepsilon \in M, \ \forall a \in M, \ \varepsilon \oplus a = a \). Although \( \oplus \) must be associative, it is not necessarily commutative, as demonstrated in structures like the \emph{bicyclic semigroup}~\cite{}\footnote{Despite its name, this is a monoid.} and Rabin-Karp's \emph{second fingerprinting method for string matching}~\cite[Section 6]{Karp:1987:ERP}. In a scan, the result at the \( n \)-th element is the reduction of the preceding subset of elements in the sequence. If the subset includes the \( n \)-th element, it is called \emph{inclusive}; if it excludes the \( n \)-th element, it is called \emph{exclusive}. The most common scan type is the prefix sum, where \( \oplus \) is addition. For example:
\[
  x = [x_1, x_2, x_3, \dots, x_n] \ \ \ \ \ \ \ y = [1, 1, 1, 1, 1]
\]
\[
  \text{InclusiveScan}(x, \oplus) = [x_1, x_1 \oplus x_2, x_1 \oplus x_2 \oplus x_3, \dots, x_1 \oplus x_2 \oplus \cdots \oplus x_n]
\]
\[
  \text{InclusiveScan}(y, +) = [1, 2, 3, 4, 5]
\]
Due to its significance in circuits and as a fundamental algorithmic primitive, scan has been extensively studied in both electrical engineering and computer science. Harris~\cite{1292373} offers a taxonomy that relates depth, fanout, and wire tracks, while Hinze~\cite{10.1007/978-3-540-27764-4_11} develops an algebraic framework for scans. Snir~\cite{10.1016/0196-67748690003-9} proved that depth $d$ and size $s$ are related by $s + d \ge 2n - 2$, and Fich~\cite{10.1145/800061.808738} proved that among minimum depth scans Ladner-Fischer~\cite{10.1145/322217.322232} networks have optimal size. Blelloch~\cite{} adapted scan to the PRAM model and popularized its use as an algorithmic primitive. Finally, Merrill and Garland~\cite{Merrill2016} provide a review of GPU scan implementations, complementing earlier work by Merrill and Grimshaw~\cite{Merrill2009}.

@@

But, while \emph{Chained-Scan} with \emph{Decoupled Lookback} appears to be the ideal scan architecture in terms of its theoretical and implementation characteristics, it falls short in meeting the criterion of \emph{portability}. This limitation arises because \emph{Chained-Scan} depends on a \emph{forward-progress guarantee}, a feature provided by NVIDIA devices but notably lacking on Apple's M-series and ARM GPUs~\cite{sorensen2}. Without this guarantee, \emph{Chained-Scan} is highly prone to catastrophic failure in the form of deadlocking.
